% =============
\section{Background}

We now provide the background for our study.
First, we depict the overall architecture of the delivery system.
Then, we recall the main geometric layouts for spherical videos.

\subsection{Navigable 360-degree Video Delivery}

The principles of a navigable video delivery system are similar as in adaptive bit-rate
video systems such as \ac{DASH}. The server offers multiple versions of the same video
and the client
selects the most appropriate version according to some criteria. These versions
are cut into second-long segments such that the client can regularly switch from one
version to another. \AD{If the client moves the head before receiving a new video version, 
he/she will be displayed a new FoV video with gradually degrading quality that is adjusted to the head movement (as described in Section \ref{sec:context}). Exactly how many seconds should pass until receiving a new video version depends on the video content, the viewer's head movements (i.e., whether he/she watches a football match, news, or plays a computer game), and the experienced quality degradations variations, which needs to be investigated in the future work.}{}

In the case of wide video with different spatial qualities, the main idea is to spatially cut
the video into \emph{tiles}.
Then, two implementations are possible for the delivery system. \AD{First option is that}{} the server
offers each tile at different qualities. The client selects each tile version independently
and it has to reconstruct
the full video from these tiles before the \ac{FoV} extraction. This solution allows a
fine setting of qualities but
most of the computation is done at the client side. The second option,
which is the one we consider in this paper and is also the industrial implementation described
by~\citet{facebook}, is that
the server prepares $x$ versions related to $x$ different \acp{QEC}. Each version
is an arrangement of tile
qualities such that the tiles that are close to the \ac{QEC} are at high-quality
and the other tiles
are at a lower quality. The main advantages include an easy management of the server
(\textit{e.g.} small \emph{manifest} file), a simple selection process for the client (by
a distance computation), and no need of re-constructing the video before the \ac{FoV} extraction.

At the client side, the end-user moves its head to decide the \ac{FoV}. The head movements
are \AD{forward and backwards, side to side, and should to shoulder, referred to as \emph{pitch}, \emph{yaw}, and \emph{roll}, respectively.}{} 
 The center of the \ac{FoV} is a
point on the sphere, the size of the \ac{FoV} depends on the device (typically
around 100$^\circ$ in state-of-the-art devices), and the orientation of the extracted video
is related to the roll.

The complete analysis and evaluation of the navigable 360-degree video delivery system
is left for future work. Due to lack of space, we focus here on the geometric layout
of the video versions and the tile quality arrangement.


\subsection{Geometric Layouts for Spherical Videos}

The projection of a sphere into a plane (known as mapping) has been extensively studied
for centuries. In this paper, we consider the four projections that are the most natural
candidates for 360-degree video delivery. These layouts are depicted in Figure~\ref{fig:mapping}.

\begin{figure}[ht]
\centering
\begin{tikzpicture}
\def\sizeSphere{20}%pt
\def\ecartY{-1.2}%cm
\def\ecartX{6}

% da sphere
\pic [local bounding box=spher]  at (0,0) {spherical=15};

% recantagular
\pic [local bounding box=equi] at (-3,\ecartY) {equirectangular={\sizeSphere}{-1}{0}};

% cupe map
\pic [local bounding box=cubemap] at (-1,\ecartY) {cubemap=\sizeSphere};

% pyramid
\pic [local bounding box=pyra] at (1,\ecartY) {pyramid=\sizeSphere};

% rhombic
%\pgfdeclareimage[width=36 pt]{dodecahedron}{RhombicDodecahedron.png}
%\node at (3,\ecartY) (dodec)
%    {\pgfbox[center,center]{\pgfuseimage{dodecahedron}}};

\def\unitused{0.22}

\pic [local bounding box=dodeca] at (3,0.88*\ecartY) {dodecahedron=\unitused};

% links
\draw[-latex] (spher.180) -| (equi);
\draw[-latex] (spher.200) -| (cubemap);
\draw[-latex] (spher.340) -| (pyra);
\draw[-latex] (spher) -| (dodeca);

\node[font=\scriptsize,anchor=north] at (equi.south) {equirectangular};
\node[font=\scriptsize,anchor=north] at (cubemap.south) {cube map};
\node[font=\scriptsize,anchor=north] at (pyra.south) {pyramid};
\node[font=\scriptsize,anchor=north] at (dodeca.south) {\vphantom{y}dodecahedron};

\end{tikzpicture}
\caption{Projections into four geometric layouts}\label{fig:mapping}
\end{figure}

The advantages and shortcomings of every projection have been studied in the literature. From
the images that are
projected on an equirectangular, a cube map, and a rhombic dodecahedron, it is possible
to generate a \ac{FoV}
for any position and angle in the sphere. Indeed, no information is loss, because no pixel is under-sampled in
this projection (no pair of pixels in the sphere is projected on these geometrical
layouts in only one pixel). However,
some pixels from the spherical image are over-sampled in the projected image. It is typically the case for
the equirectangular layout, for which the projection generates a significant over-sampling at the poles. On the
contrary, the pyramid is not a geometric layout for lossless projections. Some pixels (those who are in the back
of the view face) are under-sampled, so two pixels can be projected into one pixel by interpolating their color
values. A \ac{FoV} that is extracted for positions near the back can suffer from distortion. \AD{Hence, this is the least probable head orientation of the video viewer.}{} 